---
title: "Stage 1a: Scrape Ballotpedia recall pages (HTML cache)"
---

```{r}
#| label: purpose-block
#| include: false
# =============================================================================
# Purpose : Scrape all mayoral recall attempts from Ballotpedia (2014-2024)
# Inputs  : Ballotpedia HTML pages (fetched from web OR read from disk cache)
# Outputs : data/raw/ballotpedia/html/YEAR.html            (year summary cache)
#           data/raw/ballotpedia/html/<RecallName>.html    (individual page cache)
# =============================================================================
```

## Setup

```{r}
#| label: load-packages
library(tidyverse)
library(rvest)        # HTML parsing
library(httr2)        # HTTP requests
library(here)         # Portable file paths

here::i_am("_quarto.yml")   # Anchor here() to project root

dir.create(here("data", "raw", "ballotpedia", "html"), recursive = TRUE, showWarnings = FALSE)
```

## Step 1: robots.txt Verification

Ballotpedia's `robots.txt` does not disallow `/Political_recall_efforts,_*`, so
the recall pages may be scraped. (Verified manually.)

## Step 2: Fetch/Cache Helper

```{r}
#| label: fetch-helper
# User-Agent identifying the scraper as academic research.
ua_string <- paste0(
  "Mozilla/5.0 (compatible; AcademicResearchBot/1.0; ",
  "BPE Thesis - Transit Spending Study)"
)

# fetch_or_cache(): reads from disk if cached, otherwise fetches and saves.
fetch_or_cache <- function(url, cache_path, delay = 2) {

  if (file.exists(cache_path)) {
    message("  [CACHE] ", basename(cache_path))
    return(readLines(cache_path, warn = FALSE, encoding = "UTF-8"))
  }

  message("  [FETCH] ", url)

  # httr2 builder pattern: request → modify → perform
  resp <- request(url) |>
    req_headers(`User-Agent` = ua_string) |>
    req_options(ssl_verifypeer = 0L) |>
    req_timeout(30) |>
    req_retry(max_tries = 3) |>
    req_perform()

  # Split response body into lines before writing (preserves encoding).
  lines <- strsplit(resp_body_string(resp), "\n", fixed = TRUE)[[1]]
  writeLines(lines, cache_path)
  message("  [SAVED] ", basename(cache_path))

  Sys.sleep(delay)   # Polite delay between requests
  lines
}
```

## Step 3: Build URLs (2014-2024)

```{r}
#| label: build-urls
# URL pattern: https://ballotpedia.org/Political_recall_efforts,_YEAR
# (NOT /wiki/United_States_recall_elections,_YEAR – that returns 404)

years <- 2014:2024
base_url <- "https://ballotpedia.org/Political_recall_efforts,_"
urls <- paste0(base_url, years)

tibble(year = years, url = urls) |>
  knitr::kable(caption = "Ballotpedia summary page URLs") |>
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

## Step 4: Extract Mayoral Recall Links from Summary Pages

```{r}
#| label: extract-mayor-links
# Ballotpedia summary pages are organized as:
#   <h1><span class="mw-headline" id="Mayors/presidents">Mayors/presidents</span></h1>
#   <ul>
#     <li><a href="/Official_Name_recall,_City,_State_(Year)">...</a></li>
#     ...
#   </ul>
#   <h1><span id="School_boards">...</h1>   <- next section
#
# IMPORTANT page structure notes:
#   - Section headings are <h1> (not <h2> as you'd expect for subsections)
#   - The id is literally "Mayors/presidents" (slash is valid but unusual in IDs)
#   - Data is in <ul><li> links, NOT in <table> elements
#
# XPath selects all <a> elements in <ul> lists that follow the
# Mayors/presidents h1 heading and come before the next h1 heading.
# "following-sibling" traverses siblings in document order.

extract_mayor_links <- function(html_doc) {

  # XPath explanation (equivalent to "all <a> under <ul> after the Mayors h1"):
  #   //h1[.//span[@id='Mayors/presidents']]   = the Mayors h1
  #   /following-sibling::ul[1]                = first <ul> immediately after it
  #   //a                                      = all <a> inside that ul

  links <- html_doc |>
    html_elements(xpath = "//h1[.//span[@id='Mayors/presidents']]/following-sibling::ul[1]//a")

  if (length(links) == 0) {
    # Fallback: if XPath misses (page structure varied that year), use label filter.
    # Catches links with "mayor" or "president" in their label text or href.
    message("  [WARN] Mayors/presidents section not found via XPath – using label fallback.")
    all_links <- html_doc |> html_elements("a[href]")
    links <- all_links[
      str_detect(html_attr(all_links, "href"),  "(?i)mayor|president") |
      str_detect(html_text(all_links, trim=TRUE), "(?i)mayor|president")
    ]
  }

  tibble(
    href  = html_attr(links, "href"),
    label = html_text(links, trim = TRUE)
  ) |>
    filter(
      str_detect(href, "^/[A-Z]"),                              # Real article pages
      !str_detect(href, "Category:|Special:|Ballotpedia:|#"),   # Not meta pages
      str_detect(href, "(?i)recall")                            # Must be a recall page
    ) |>
    distinct(href, .keep_all = TRUE)
}

# ── Apply to all years ────────────────────────────────────────────────────
message("Extracting mayoral recall links for ", length(years), " years...")

links_by_year <- purrr::map2(
  years, urls,
  function(yr, url) {
    message("Year ", yr, ":")
    cache_path <- here("data", "raw", "ballotpedia", "html", paste0(yr, ".html"))
    html_lines <- fetch_or_cache(url, cache_path, delay = 2)
    html_doc   <- read_html(paste(html_lines, collapse = "\n"))
    links      <- extract_mayor_links(html_doc)
    links |> mutate(year = yr)
  }
)
names(links_by_year) <- as.character(years)

# Stack all years' links.
all_mayor_links <- bind_rows(links_by_year)

cat("\n=== Links per year ===\n")
all_mayor_links |>
  count(year, name = "n_links") |>
  mutate(flag = ifelse(n_links == 0, "WARNING: Zero links – check page structure", "OK")) |>
  knitr::kable(caption = "Mayoral recall links found per year") |>
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

## Step 5: Scrape Individual Recall Pages

```{r}
#| label: scrape-individual-pages
# For each individual recall page, extract:
# (A) Structured fields from the URL: official name, city, state, filing year
#     URL format: /Official_Name_recall,_City,_State_(Year)
# (B) Outcome from wgCategories JSON embedded in the page's <script> block
#
# wgCategories is the cleanest outcome source: the MediaWiki page category
# list, which Ballotpedia maintains consistently. The category names map to
# the four outcomes:
#   "Recall, did not go to a vote" → Did Not Make Ballot
#   "Recall won"                    → Removed
#   "Recall lost"                   → Defeated
#   (resigned before vote)          → Resigned  (inferred from category text)

# Map Ballotpedia categories to outcome codes.
parse_outcome_from_categories <- function(cats) {
  dplyr::case_when(
    any(str_detect(cats, "(?i)recalled and removed|Recall won|removed from office")) ~ "Removed",
    any(str_detect(cats, "(?i)resigned"))                                             ~ "Resigned",
    any(str_detect(cats, "(?i)Recall lost|recall defeated|not removed"))              ~ "Defeated",
    any(str_detect(cats, "(?i)did not go to a vote"))                                 ~ "Did Not Make Ballot",
    any(str_detect(cats, "(?i)Unsuccessful"))                                         ~ "Did Not Make Ballot",
    TRUE                                                                               ~ "Unknown"
  )
}

# Scrape one individual recall page.
scrape_individual_page <- function(href, delay = 1.5) {

  # Build a filesystem-safe cache filename from the URL path.
  safe_name  <- str_replace_all(basename(href), "[^A-Za-z0-9._-]", "_")
  cache_path <- here("data", "raw", "ballotpedia", "html", paste0(safe_name, ".html"))

  lines <- fetch_or_cache(paste0("https://ballotpedia.org", href), cache_path, delay = delay)
  src   <- paste(lines, collapse = " ")   # Flat string for regex

  # ── Extract wgCategories from embedded JS ─────────────────────────────
  # The page contains a JS block like:
  #   RLCONF={"wgCategories":["Recall, Missouri","Recall, 2020", ...], ...}
  # Extract the category array with a regex.
  cats_match <- str_extract(src, '"wgCategories":\\[[^\\]]+\\]')
  cats <- if (!is.na(cats_match)) {
    str_extract_all(cats_match, '"([^"]+)"')[[1]] |>
      str_remove_all('"') |>
      (\(x) x[x != "wgCategories"])()   # Remove the key name itself
  } else {
    character(0)
  }

  # ── Parse URL for structured fields ───────────────────────────────────
  # URL format examples:
  #   /Quinton_Lucas_recall,_Kansas_City,_Missouri_(2020)
  #   /Mayor_and_city_council_recall,_Westminster,_California_(2020)
  url_body     <- str_remove(href, "^/")
  official_raw <- url_body |>
    str_extract("^[^,]+") |>
    str_remove("_recall$") |>
    str_replace_all("_", " ")
  after_comma   <- str_extract(url_body, "(?<=recall,_).+$")
  year_raw      <- str_extract(after_comma, "(?<=\\()([0-9]{4})")
  location_part <- str_remove(after_comma, "_\\([0-9-]+\\).*$")
  loc_parts     <- str_split(location_part, ",_")[[1]]
  city_raw      <- str_replace_all(loc_parts[1], "_", " ")
  state_raw     <- if (length(loc_parts) >= 2)
    str_replace_all(loc_parts[length(loc_parts)], "_", " ")
  else
    NA_character_

  tibble(
    href        = href,
    official    = official_raw,
    city        = city_raw,
    state       = state_raw,
    filing_year = as.integer(year_raw),
    outcome     = parse_outcome_from_categories(cats),
    categories  = paste(cats, collapse = "; ")
  )
}

# ── Apply to all links ─────────────────────────────────────────────────────
message("Scraping ", nrow(all_mayor_links), " individual recall pages...")
message("(2-second delay between requests; estimated time: ",
        round(nrow(all_mayor_links) * 2 / 60, 1), " minutes)")

recalls_raw <- purrr::map_dfr(
  all_mayor_links$href,
  scrape_individual_page,
  delay = 2   # 2-second delay between individual page requests (polite)
)
```

## Step 6: Audit Checks

### 6a – Record Count Per Year

```{r}
#| label: audit-count-per-year
count_by_year <- recalls_raw |>
  count(filing_year, name = "n_records") |>
  mutate(
    flag = case_when(
      n_records == 0  ~ "ERROR: Zero records",
      n_records < 3   ~ "WARNING: Very few records",
      TRUE            ~ "OK"
    )
  )

count_by_year |>
  knitr::kable(caption = "Mayoral recall records per year") |>
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

### 6b – Outcome Distribution

```{r}
#| label: audit-outcomes
recalls_raw |>
  count(outcome, sort = TRUE) |>
  mutate(pct = scales::percent(n / sum(n), accuracy = 0.1)) |>
  knitr::kable(caption = "Recall outcome distribution (all years)") |>
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

### 6c – Spot-Check Known Recalls

```{r}
#| label: audit-spotcheck
# Verify that known recalls appear correctly.
known <- tribble(
  ~city,         ~state,       ~year, ~expected_outcome,      ~note,
  "Kansas City", "Missouri",   2020,  "Did Not Make Ballot",  "Quinton Lucas – didn't get enough signatures",
  "Los Angeles", "California", 2020,  "Did Not Make Ballot",  "Eric Garcetti – did not advance",
  "Seattle",     "Washington", 2020,  "Did Not Make Ballot",  "Jenny Durkan 2020 attempt"
)

for (i in seq_len(nrow(known))) {
  r   <- known[i, ]
  hit <- recalls_raw |>
    filter(str_detect(tolower(city), tolower(r$city)),
           str_detect(state, r$state),
           filing_year == r$year)
  status <- if (nrow(hit) > 0) {
    paste("FOUND – outcome:", hit$outcome[1],
          if (hit$outcome[1] == r$expected_outcome) "[CORRECT]" else "[MISMATCH]")
  } else "NOT FOUND"
  cat(sprintf("  %s | %s, %s (%d): %s\n", status, r$city, r$state, r$year, r$note))
}
```

### 6d – "Unknown" Outcomes

```{r}
#| label: audit-unknown-outcomes
# Flag any records where the outcome couldn't be determined.
# These need manual review – either the category regex needs updating,
# or the Ballotpedia page uses non-standard categories.
unknown_outcomes <- recalls_raw |> filter(outcome == "Unknown")
cat("Records with Unknown outcome:", nrow(unknown_outcomes), "\n")
if (nrow(unknown_outcomes) > 0) {
  unknown_outcomes |>
    select(official, city, state, filing_year, categories) |>
    knitr::kable(caption = "Unknown outcomes – review categories to improve classifier") |>
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
}
```

## Step 7: In-memory summary (no disk write)

```{r}
#| label: save-output
# Scrape-only: this stage caches HTML; it writes no .rds.
cat("In-memory recall records (this session):", nrow(recalls_raw), "\n")
cat("Columns:", paste(names(recalls_raw), collapse = ", "), "\n")
glimpse(recalls_raw)
```
