---
title: "Stage 6: Exploratory Data Analysis"
---

```{r}
#| label: purpose
#| include: false
# =============================================================================
# Purpose : Descriptive look at panel.rds before modeling. Coverage heatmap,
#           outcome distributions, treated-vs-control time series, pooled
#           pre-trend visualization. No causal inference here.
# Inputs  : data/final/panel.rds
# Outputs : Inline HTML; figures saved to output/figures/.
# =============================================================================
```

## Setup

```{r}
#| label: setup
library(tidyverse)
library(here)
library(kableExtra)
library(patchwork)
here::i_am("_quarto.yml")

panel <- readRDS(here("data", "final", "panel.rds"))

dir.create(here("output", "figures"), recursive = TRUE, showWarnings = FALSE)
dir.create(here("output", "tables"), recursive = TRUE, showWarnings = FALSE)

theme_set(theme_minimal(base_size = 12) +
          theme(plot.title    = element_text(face = "bold"),
                plot.subtitle = element_text(color = "grey40")))

palette_ab <- c("treated" = "#c0392b", "control" = "#2980b9")

cat(sprintf("Panel: %d rows · %d cities · years %d–%d\n",
            nrow(panel), n_distinct(panel$city_id),
            min(panel$year), max(panel$year)))
```

---

## Geographic distribution of recall petitions

```{r}
#| label: fig-recall-heatmap
#| fig-cap: "Municipal recall petitions, 2014-2024, by county. Counties with no recall activity are unshaded. AK and HI shifted to insets."
#| fig-width: 11
#| fig-height: 6
suppressPackageStartupMessages({
  library(sf)
  library(tigris)
})
options(tigris_use_cache = TRUE, tigris_class = "sf")
source(here("R", "std_city.R"))

recalls <- readRDS(here("data", "cleaned", "recalls_clean.rds"))

# City -> county FIPS via the same Place + Cousub union 03c uses.
state_abb <- tibble(abb = state.abb, name = tolower(state.name)) |>
  add_row(abb = "DC", name = "district of columbia")

build_xwalk <- function(path, name_col) {
  readr::read_delim(path, delim = "|", show_col_types = FALSE) |>
    janitor::clean_names() |>
    left_join(state_abb, by = c("state" = "abb")) |>
    transmute(
      place_norm = std_city(.data[[name_col]]),
      state_norm = coalesce(name, tolower(state)),
      fips_county = paste0(statefp, countyfp)
    )
}

pc <- build_xwalk(here("data","raw","census","national_place_by_county2020.txt"), "placename")
cs <- build_xwalk(here("data","raw","census","national_cousub2020.txt"),         "cousubname")

recalls_key <- recalls |>
  mutate(city_norm  = std_city(city_std),
         state_norm = str_to_lower(state_std))

# Resolve each petition to a county: Place file first, Cousub fallback for
# MCD-style municipalities (NJ townships, MI charter townships, etc.).
matched_place <- recalls_key |>
  inner_join(pc, by = c("city_norm" = "place_norm", "state_norm"),
             relationship = "many-to-many")
matched_cousub <- recalls_key |>
  anti_join(pc, by = c("city_norm" = "place_norm", "state_norm")) |>
  inner_join(cs, by = c("city_norm" = "place_norm", "state_norm"),
             relationship = "many-to-many")

petitions_by_county <- bind_rows(matched_place, matched_cousub) |>
  count(fips_county, name = "n_petitions") |>
  mutate(state_fp = substr(fips_county, 1, 2))

n_excluded_unmatched <- nrow(recalls_key) - sum(petitions_by_county$n_petitions)
n_excluded_akhi      <- sum(petitions_by_county$n_petitions[
                              petitions_by_county$state_fp %in% c("02","15")])

cat("Petitions resolved to a county:", sum(petitions_by_county$n_petitions),
    "of", nrow(recalls_key), "\n")
cat("Counties with at least one petition:", nrow(petitions_by_county), "\n")

non_conus <- c("AK","HI","AS","GU","MP","PR","VI")

counties_sf <- counties(cb = TRUE, year = 2020, progress_bar = FALSE) |>
  filter(!STATE_NAME %in% c("Alaska","Hawaii","American Samoa","Guam",
                            "Commonwealth of the Northern Mariana Islands",
                            "Puerto Rico","United States Virgin Islands")) |>
  mutate(fips_county = paste0(STATEFP, COUNTYFP)) |>
  left_join(petitions_by_county, by = "fips_county")

states_sf <- states(cb = TRUE, year = 2020, progress_bar = FALSE) |>
  filter(!STUSPS %in% non_conus)

ggplot() +
  geom_sf(data = counties_sf, aes(fill = n_petitions),
          color = "white", linewidth = 0.05) +
  geom_sf(data = states_sf, fill = NA, color = "grey30", linewidth = 0.3) +
  scale_fill_distiller(palette = "Reds", direction = 1, na.value = "grey92",
                       name = "Petitions\nfiled",
                       breaks = scales::pretty_breaks(4)) +
  coord_sf(crs = 5070) +
  labs(title    = "Municipal recall petitions, 2014-2024",
       subtitle = "County-level counts, contiguous 48 states",
       caption  = sprintf(paste0("Source: Ballotpedia Recall Reports. City→county via Census 2020 Place + Cousub crosswalks. ",
                                 "Excluded: %d petitions in AK/HI, %d in cities not matched to a county."),
                          n_excluded_akhi, n_excluded_unmatched)) +
  theme_void(base_size = 12) +
  theme(plot.title    = element_text(face = "bold"),
        plot.subtitle = element_text(color = "grey40"),
        legend.position = "right")

ggsave(here("output", "figures", "fig_recall_heatmap.png"),
       width = 11, height = 6, dpi = 150)
```

---

## 1 – Sample membership

```{r}
#| label: tab-samples
panel |>
  summarise(
    `Model A HQ (cities)`  = n_distinct(city_id[in_modelA_hq]),
    `Model A ext (cities)` = n_distinct(city_id[in_modelA_ext]),
    `Model B HQ (cities)`  = n_distinct(city_id[in_modelB_hq]),
    `Model B ext (cities)` = n_distinct(city_id[in_modelB_ext]),
    `Treated A HQ`         = n_distinct(city_id[in_modelA_hq & group_A == "treated"]),
    `Control A HQ`         = n_distinct(city_id[in_modelA_hq & group_A == "control"])
  ) |>
  pivot_longer(everything(), names_to = "Sample", values_to = "N") |>
  kable(caption = "City counts by sample membership") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r}
#| label: tab-rows-after-na
# Effective regression sample sizes after NA drop on the outcome.
panel |>
  summarise(
    `Model A HQ (agency-city-year)` = sum(in_modelA_hq & !is.na(log_transit_cap)),
    `Model B HQ (agency-city-year)` = sum(in_modelB_hq & !is.na(log_transit_cap)),
    `Model A ext`                   = sum(in_modelA_ext & !is.na(log_transit_cap)),
    `Model B ext`                   = sum(in_modelB_ext & !is.na(log_transit_cap))
  ) |>
  pivot_longer(everything(), names_to = "Sample", values_to = "City-year rows") |>
  kable(caption = "Effective regression rows (after NA drop on log_transit_cap)") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r}
#| label: tab-outcome-cat
panel |>
  filter(group_A == "treated") |>
  distinct(city_id, outcome_cat) |>
  count(outcome_cat, name = "cities") |>
  kable(caption = "Treated-city composition by outcome category") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## 2 – NTD coverage diagnostic

```{r}
#| label: fig-coverage-heatmap
#| fig-cap: "Share of cities with an NTD report each year, by Model A group"
#| fig-height: 4

cov <- panel |>
  filter(in_modelA_hq, !is.na(group_A)) |>
  group_by(group_A, year) |>
  summarise(
    n_cities    = n_distinct(city_id),
    n_observed  = n_distinct(city_id[ntd_observed]),
    share_obs   = n_observed / n_cities,
    .groups = "drop"
  )

ggplot(cov, aes(x = factor(year), y = group_A, fill = share_obs)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%d/%d", n_observed, n_cities)),
            size = 3, color = "white") +
  scale_fill_gradient(low = "#7f8c8d", high = "#27ae60",
                      labels = scales::percent, limits = c(0, 1),
                      name = "Share with\nNTD report") +
  labs(title = "NTD reporting coverage by year and group",
       subtitle = "Cells: cities-with-report / cities-in-group. Model A HQ sample.",
       x = "Year", y = NULL)

ggsave(here("output", "figures", "fig_coverage_heatmap.png"),
       width = 9, height = 3.5, dpi = 150)
```

```{r}
#| label: fig-coverage-by-city
#| fig-cap: "Per-city NTD coverage timeline (Model A sample)"
#| fig-height: 9

city_obs <- panel |>
  filter(in_modelA_hq) |>
  mutate(grp = ifelse(city_id %in% panel$city_id[panel$group_A == "treated"],
                      "treated", "control"))

# Order cities: treated first, then by total years observed
city_order <- city_obs |>
  group_by(city_id, grp) |>
  summarise(n_obs = sum(ntd_observed), .groups = "drop") |>
  arrange(grp, desc(n_obs)) |>
  pull(city_id)

city_obs |>
  mutate(city_id = factor(city_id, levels = city_order)) |>
  ggplot(aes(x = year, y = city_id, fill = ntd_observed)) +
  geom_tile(color = "white") +
  scale_fill_manual(values = c(`TRUE` = "#27ae60", `FALSE` = "#ecf0f1"),
                    name = "NTD observed") +
  scale_x_continuous(breaks = 2014:2024) +
  facet_grid(grp ~ ., scales = "free_y", space = "free_y") +
  labs(title = "NTD reporting timeline by city",
       subtitle = "Green = NTD report exists for that city-year. Sorted by total years observed within group.",
       x = "Year", y = NULL) +
  theme(axis.text.y = element_text(size = 6))

ggsave(here("output", "figures", "fig_coverage_by_city.png"),
       width = 9, height = 11, dpi = 150)
```

---

## 3 – Outcome distribution

```{r}
#| label: fig-outcome-dist
#| fig-cap: "Distribution of total transit capital spending (observed city-years)"
#| fig-height: 4

obs <- panel |> filter(ntd_observed)

p_raw <- ggplot(obs |> filter(total_capital_usd > 0),
                aes(x = total_capital_usd)) +
  geom_histogram(bins = 50, fill = "#2980b9", alpha = 0.85) +
  scale_x_log10(labels = scales::dollar_format(scale = 1e-6, suffix = "M",
                                                accuracy = 0.1)) +
  labs(title = "Capital spending (log dollar axis)",
       x = "Total capital, USD (log scale)", y = "Agency-city-years")

p_log <- ggplot(obs, aes(x = log_transit_cap)) +
  geom_histogram(bins = 50, fill = "#27ae60", alpha = 0.85) +
  labs(title = "log(transit capital + 1) – regression DV",
       x = "log(transit capital + 1)", y = "Agency-city-years")

p_raw + p_log +
  plot_annotation(subtitle = sprintf("N = %d agency-city-years with NTD report", nrow(obs)))

ggsave(here("output", "figures", "fig_outcome_dist.png"),
       width = 10, height = 4, dpi = 150)
```

```{r}
#| label: tab-outcome-summary
panel |>
  filter(in_modelA_hq, !is.na(group_A), ntd_observed) |>
  group_by(group_A) |>
  summarise(
    `City-years (observed)` = n(),
    `Median capital ($M)`   = round(median(total_capital_usd) / 1e6, 2),
    `Mean log_transit_cap`  = round(mean(log_transit_cap), 2),
    `SD log_transit_cap`    = round(sd(log_transit_cap), 2),
    .groups = "drop"
  ) |>
  kable(caption = "Outcome summary by Model A group (observed years only)") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## 4 – Transit capital over time

### 4a – By coverage tier (recall-active population)

```{r}
#| label: fig-capital-by-tier
#| fig-cap: "Mean log transit capital by coverage tier (EXT/HQ/REST), recall-active population, by year"
palette_tier <- c("EXT" = "#16a085", "HQ" = "#c0392b", "REST" = "#8e44ad")

tier_ts <- bind_rows(
  panel |> filter(in_modelA_ext)  |> mutate(tier = "EXT"),
  panel |> filter(in_modelA_hq)   |> mutate(tier = "HQ"),
  panel |> filter(in_modelA_rest) |> mutate(tier = "REST")
) |>
  filter(!is.na(log_transit_cap)) |>
  group_by(tier, year) |>
  summarise(
    n_obs  = n(),
    mean_y = mean(log_transit_cap),
    se     = sd(log_transit_cap) / sqrt(n_obs),
    ci_lo  = mean_y - 1.96 * se,
    ci_hi  = mean_y + 1.96 * se,
    .groups = "drop"
  ) |>
  mutate(tier = factor(tier, levels = c("EXT", "HQ", "REST")))

ggplot(tier_ts, aes(x = year, y = mean_y, color = tier, fill = tier)) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.12, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(aes(size = n_obs)) +
  scale_color_manual(values = palette_tier) +
  scale_fill_manual(values = palette_tier) +
  scale_x_continuous(breaks = 2014:2024) +
  scale_size_continuous(range = c(1.5, 4.5), name = "N obs") +
  labs(title = "Mean log(transit capital) by coverage tier and calendar year",
       subtitle = "Full recall-active population (treated + control). Shaded band: 95% CI. Tiers nested: REST within HQ within EXT.",
       x = "Year", y = "Mean log(transit capital + 1)",
       color = "Tier", fill = "Tier")

ggsave(here("output", "figures", "fig_capital_by_tier.png"),
       width = 9, height = 5, dpi = 150)
```

### 4b – Treated vs. control (Model A sample)

```{r}
#| label: fig-time-series
#| fig-cap: "Mean log transit capital, treated vs. control, by year (Model A sample)"

ts <- panel |>
  filter(in_modelA_hq, !is.na(group_A), !is.na(log_transit_cap)) |>
  group_by(group_A, year) |>
  summarise(
    n_obs   = n(),
    mean_y  = mean(log_transit_cap),
    se      = sd(log_transit_cap) / sqrt(n_obs),
    ci_lo   = mean_y - 1.96 * se,
    ci_hi   = mean_y + 1.96 * se,
    .groups = "drop"
  )

ggplot(ts, aes(x = year, y = mean_y, color = group_A, fill = group_A)) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.18, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(aes(size = n_obs)) +
  scale_color_manual(values = palette_ab) +
  scale_fill_manual(values = palette_ab) +
  scale_x_continuous(breaks = 2014:2024) +
  scale_size_continuous(range = c(1.5, 4.5), name = "N obs") +
  labs(title = "Mean log(capital) by group, calendar year",
       subtitle = "Shaded band: 95% CI from group-year mean. Model A sample, observed city-years only.",
       x = "Year", y = "Mean log_transit_cap",
       color = "Group", fill = "Group")

ggsave(here("output", "figures", "fig_time_series.png"),
       width = 9, height = 5, dpi = 150)
```

---

## 5 – Pre-trend visualization (event time)

```{r}
#| label: fig-event-coverage
#| fig-cap: "Per-treated-city NTD coverage in event time"
#| fig-height: 4

et_cov <- panel |>
  filter(in_modelA_hq, group_A == "treated", !is.na(event_time_A)) |>
  mutate(city_id = fct_reorder(city_id, first_success_year))

ggplot(et_cov, aes(x = event_time_A, y = city_id, fill = ntd_observed)) +
  geom_tile(color = "white") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey30") +
  scale_fill_manual(values = c(`TRUE` = "#27ae60", `FALSE` = "#ecf0f1"),
                    name = "NTD observed") +
  scale_x_continuous(breaks = seq(-10, 10, 2)) +
  labs(title = "Event-time coverage, treated cities",
       subtitle = "Each row = one treated city. Dashed line = recall year.",
       x = "Event time (years from first successful recall)", y = NULL) +
  theme(axis.text.y = element_text(size = 8))

ggsave(here("output", "figures", "fig_event_coverage.png"),
       width = 9, height = 4, dpi = 150)
```

```{r}
#| label: fig-pretrend
#| fig-cap: "Mean log transit capital by event time (pooled Removed + Resigned)"

et <- panel |>
  filter(in_modelA_hq,
         group_A == "treated",
         !is.na(event_time_A),
         !is.na(log_transit_cap),
         event_time_A >= -5, event_time_A <= 5) |>
  group_by(event_time_A) |>
  summarise(
    n_obs   = n(),
    n_city  = n_distinct(city_id),
    mean_y  = mean(log_transit_cap),
    se      = sd(log_transit_cap) / sqrt(n_obs),
    ci_lo   = mean_y - 1.96 * se,
    ci_hi   = mean_y + 1.96 * se,
    .groups = "drop"
  )

ggplot(et, aes(x = event_time_A, y = mean_y)) +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_hline(yintercept = mean(et$mean_y[et$event_time_A < 0]),
             linetype = "dotted", color = "grey60") +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#c0392b", alpha = 0.18) +
  geom_line(color = "#c0392b", linewidth = 0.9) +
  geom_point(aes(size = n_city), color = "#c0392b") +
  scale_x_continuous(breaks = -5:5,
                     labels = function(x) ifelse(x == 0, "Recall\nyear",
                                                  as.character(x))) +
  scale_size_continuous(range = c(2, 6), name = "N treated\ncities") +
  labs(title = "Raw event-time means – treated cities only",
       subtitle = "Pooled Removed + Resigned. ±5 years. Descriptive – no controls or FE.",
       x = "Years relative to first successful recall",
       y = "Mean log_transit_cap")

ggsave(here("output", "figures", "fig_pretrend.png"),
       width = 9, height = 5, dpi = 150)
```

```{r}
#| label: tab-event-time
et |>
  select(event_time_A, n_city, n_obs, mean_y, ci_lo, ci_hi) |>
  mutate(across(c(mean_y, ci_lo, ci_hi), ~ round(.x, 2))) |>
  kable(caption = "Event-time bins – treated cities",
        col.names = c("event time", "N cities", "N obs", "mean", "95% lo", "95% hi")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## 6 – Recall activity by year

```{r}
#| label: fig-petitions-by-year
#| fig-cap: "Petition-filing years across the panel (Model B sample)"
#| fig-height: 3.5

panel |>
  filter(in_modelB_hq, any_petition_this_year) |>
  count(year) |>
  ggplot(aes(x = year, y = n)) +
  geom_col(fill = "#8e44ad", alpha = 0.85, width = 0.7) +
  geom_text(aes(label = n), vjust = -0.4, size = 3.5) +
  scale_x_continuous(breaks = 2014:2024) +
  labs(title = "Petition-filing agency-city-years per calendar year",
       subtitle = "Model B HQ sample. Each agency-city-year with a petition filed that year.",
       x = "Year", y = "Agency-city-years with petition")

ggsave(here("output", "figures", "fig_petitions_by_year.png"),
       width = 9, height = 3.5, dpi = 150)
```

---

## 7 – Control covariate distributions

```{r}
#| label: fig-controls
#| fig-cap: "Distributions of time-varying controls"
#| fig-height: 4

ctrl <- panel |> filter(in_modelB_hq)

p_unemp <- ggplot(ctrl |> filter(!is.na(unemp_rate)), aes(x = unemp_rate)) +
  geom_histogram(bins = 40, fill = "#8e44ad", alpha = 0.85) +
  labs(title = "Unemployment rate (county, BLS LAUS)",
       x = "Unemployment rate (%)", y = "City-years")

p_density <- ggplot(ctrl |> filter(!is.na(pop_density), pop_density > 0),
                    aes(x = pop_density)) +
  geom_histogram(bins = 40, fill = "#e67e22", alpha = 0.85) +
  scale_x_log10(labels = scales::comma) +
  labs(title = "Population density (Census PEP)",
       x = "People per sq mi (log scale)", y = "City-years")

p_unemp + p_density +
  plot_annotation(subtitle = "Model B sample, all city-years")

ggsave(here("output", "figures", "fig_controls.png"),
       width = 10, height = 4, dpi = 150)
```

```{r}
#| label: tab-control-summary
ctrl |>
  summarise(
    `Pop density (median)`  = round(median(pop_density, na.rm = TRUE)),
    `Pop density (p10)`     = round(quantile(pop_density, 0.10, na.rm = TRUE)),
    `Pop density (p90)`     = round(quantile(pop_density, 0.90, na.rm = TRUE)),
    `Unemp (median, %)`     = round(median(unemp_rate, na.rm = TRUE), 2),
    `Unemp (p10, %)`        = round(quantile(unemp_rate, 0.10, na.rm = TRUE), 2),
    `Unemp (p90, %)`        = round(quantile(unemp_rate, 0.90, na.rm = TRUE), 2)
  ) |>
  pivot_longer(everything(), names_to = "Statistic", values_to = "Value") |>
  kable(caption = "Control variable summary (Model B sample)") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## 8 – Missingness diagnostic

```{r}
#| label: fig-missingness
#| fig-cap: "Variable missingness by year (Model B sample)"
#| fig-height: 4

panel |>
  filter(in_modelB_hq) |>
  group_by(year) |>
  summarise(
    `log_transit_cap`        = mean(is.na(log_transit_cap)),
    `pop_density`            = mean(is.na(pop_density)),
    `unemp_rate`             = mean(is.na(unemp_rate)),
    .groups = "drop"
  ) |>
  pivot_longer(-year, names_to = "variable", values_to = "pct_missing") |>
  ggplot(aes(x = year, y = variable, fill = pct_missing)) +
  geom_tile(color = "white") +
  geom_text(aes(label = scales::percent(pct_missing, accuracy = 1)),
            size = 3, color = "white") +
  scale_fill_gradient(low = "#27ae60", high = "#c0392b",
                      labels = scales::percent, limits = c(0, 1),
                      name = "% Missing") +
  scale_x_continuous(breaks = 2014:2024) +
  labs(title = "Missingness by variable × year",
       subtitle = "Model B sample.",
       x = "Year", y = NULL)

ggsave(here("output", "figures", "fig_missingness.png"),
       width = 9, height = 3.5, dpi = 150)
```

---

## 9 – Sample cascade and statistics

### 9.1 Sample cascade

```{r}
#| label: tab-cascade

# Universe and successful-recall pool cover all recall-active cities (from
# `recalls`). The per-model rows count what enters each regression: cities,
# agency-city pairs, and petition agency-city-years remaining after dropping rows
# with missing DV or controls.
estim <- panel |>
  filter(!is.na(log_transit_cap), !is.na(log_pop_density), !is.na(unemp_rate))

get_cities <- function(flag, group = NULL) {
  keep <- estim[[flag]]
  if (!is.null(group)) keep <- keep & estim$group_A == group
  n_distinct(estim$city_id[keep])
}
get_pairs <- function(flag) n_distinct(estim$agency_city[estim[[flag]]])

n_universe <- n_distinct(recalls$city_id)
n_success  <- n_distinct(recalls$city_id[recalls$any_removed | recalls$any_resigned])

cascade_table <- tribble(
  ~Step, ~EXT, ~HQ, ~REST,
  "Recall-active universe (cities)", n_universe, n_universe, n_universe,
  "Successful recall pool (cities)", n_success, n_success, n_success,
  "Model A: treated cities (in regression)",
    get_cities("in_modelA_ext", "treated"), get_cities("in_modelA_hq", "treated"), get_cities("in_modelA_rest", "treated"),
  "Model A: control cities (in regression)",
    get_cities("in_modelA_ext", "control"), get_cities("in_modelA_hq", "control"), get_cities("in_modelA_rest", "control"),
  "Model A: agency-city pairs (in regression)",
    get_pairs("in_modelA_ext"), get_pairs("in_modelA_hq"), get_pairs("in_modelA_rest"),
  "Model B: cities (in regression)",
    get_cities("in_modelB_ext"), get_cities("in_modelB_hq"), get_cities("in_modelB_rest"),
  "Model B: agency-city pairs (in regression)",
    get_pairs("in_modelB_ext"), get_pairs("in_modelB_hq"), get_pairs("in_modelB_rest"),
  "Model B: petition agency-city-years (in regression)",
    sum(estim$in_modelB_ext & estim$any_petition_this_year),
    sum(estim$in_modelB_hq & estim$any_petition_this_year),
    sum(estim$in_modelB_rest & estim$any_petition_this_year)
)

cascade_title <- "Sample cascade"
cascade_note  <- "The first two rows cover all recall-active cities; the per-model rows count the cities, agency-city pairs, and petition agency-city-years entering each regression after dropping rows with missing controls. EXT = agency matched by headquarters, service-area place, or service-area county; HQ = agency headquartered in the city; REST = agency both headquartered in and serving the city."

writeLines(
  as.character(
    kbl(cascade_table, format = "latex", booktabs = TRUE, caption = cascade_title) |>
      footnote(general = cascade_note, general_title = "",
               footnote_as_chunk = TRUE, threeparttable = TRUE, escape = FALSE)),
  here("output", "tables", "tab_cascade.tex"))

cascade_table |>
  kable(caption = cascade_title) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  footnote(general = cascade_note, general_title = "", footnote_as_chunk = TRUE)
```

### 9.2 Sample statistics

```{r}
#| label: tab-sample-stats

stats_vars <- c("log_transit_cap", "total_capital_usd",
                "pop_density", "unemp_rate", "any_petition_this_year")

stat_labels <- c(log_transit_cap        = "log_transit_cap (DV)",
                 total_capital_usd      = "Capital, USD (raw DV)",
                 pop_density            = "Population density",
                 unemp_rate             = "Unemployment rate (%)",
                 any_petition_this_year = "Petition filed (0/1)")

summarise_sample <- function(df, vars) {
  df |>
    select(all_of(vars)) |>
    mutate(across(everything(), as.numeric)) |>
    pivot_longer(everything(), names_to = "variable") |>
    group_by(variable) |>
    summarise(
      N      = sum(!is.na(value)),
      Mean   = mean(value,   na.rm = TRUE),
      Median = median(value, na.rm = TRUE),
      SD     = sd(value,     na.rm = TRUE),
      Min    = min(value,    na.rm = TRUE),
      Max    = max(value,    na.rm = TRUE),
      .groups = "drop"
    )
}

stats_table <- function(panel, model_flags, model_label, tex_path = NULL) {
  tbl <- imap_dfr(model_flags, \(flag, label)
    panel |> filter(.data[[flag]]) |>
      summarise_sample(stats_vars) |>
      mutate(Sample = label)
  ) |>
    mutate(variable = stat_labels[variable],
           variable = factor(variable, levels = stat_labels[stats_vars]),
           Sample   = factor(Sample, levels = names(model_flags))) |>
    arrange(Sample, variable) |>
    mutate(across(c(Mean, Median, SD, Min, Max), \(x) signif(x, 3))) |>
    select(Sample, Variable = variable, N, Mean, Median, SD, Min, Max)

  cap_title <- paste0("Sample statistics: ", model_label)
  cap_note  <- "EXT = agency matched by headquarters, service-area place, or service-area county; HQ = agency headquartered in the city; REST = agency both headquartered in and serving the city."

  if (!is.null(tex_path)) {
    writeLines(
      as.character(
        kbl(tbl, format = "latex", booktabs = TRUE, caption = cap_title) |>
          footnote(general = cap_note, general_title = "",
                   footnote_as_chunk = TRUE, threeparttable = TRUE, escape = FALSE)),
      tex_path)
  }

  tbl |>
    kable(caption = cap_title) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
    footnote(general = cap_note, general_title = "", footnote_as_chunk = TRUE)
}

modelA_flags <- c("EXT" = "in_modelA_ext", "HQ" = "in_modelA_hq", "REST" = "in_modelA_rest")
modelB_flags <- c("EXT" = "in_modelB_ext", "HQ" = "in_modelB_hq", "REST" = "in_modelB_rest")

stats_table(panel, modelA_flags, "Model A", here("output", "tables", "tab_sample_stats_A.tex"))
stats_table(panel, modelB_flags, "Model B", here("output", "tables", "tab_sample_stats_B.tex"))
```

