---
title: "Stage 8: Extended models"
execute:
  cache: false
---

```{r}
#| label: purpose
#| include: false
# =============================================================================
# Purpose : Extensions to the headline models.
#   8d. Extended-coverage Model A – event-study on the in_modelA_ext sample
#       (HQ ∪ SA-place ∪ SA-county matched (city, agency) pairs).
#   8e. Extended-coverage Model B on in_modelB_ext: baseline and × covid_period.
#   8f. Restricted-coverage Model A on in_modelA_rest (HQ ∩ area-served pairs):
#       no-controls and with-controls.
#   8g. Restricted-coverage Model B on in_modelB_rest: no-controls,
#       with-controls, and × covid_period.
# All specs use the agency-city-year panel: FE | agency_city + year |, cluster
# on city_id.
# =============================================================================
```

## Setup

```{r}
#| label: setup
library(tidyverse)
library(fixest)
library(kableExtra)
library(here)
here::i_am("_quarto.yml")
source(here("R", "model_helpers.R"))

panel <- readRDS(here("data", "final", "panel.rds"))

dir.create(here("output", "tables"),  recursive = TRUE, showWarnings = FALSE)
dir.create(here("output", "figures"), recursive = TRUE, showWarnings = FALSE)

theme_set(theme_minimal(base_size = 12) +
          theme(plot.title = element_text(face = "bold"),
                plot.subtitle = element_text(color = "grey40")))

setFixest_dict(c(
  log_transit_cap          = "log(transit capital + 1)",
  log_pop_density          = "log(pop density)",
  unemp_rate               = "Unemployment rate (%)",
  any_petition_this_year   = "Petition (any)",
  agency_city              = "Agency-City",
  year                     = "Year"
))
```

---

## Extended-coverage Model A

Same event-study spec as the headline on the extended-coverage sample (`in_modelA_ext`),
which includes (city, agency) pairs matched via SA-place or SA-county in
addition to HQ.

```{r}
#| label: 8d-fit
mA_ext <- fit_modelA(panel |> filter(in_modelA_ext))

etable(mA_ext,
       title  = "Model A – extended-coverage sample",
       digits = "r3", se.below = TRUE)
```

```{r}
#| label: 8d-plot
#| fig-cap: "Model A event-study coefficients with 95% CIs – extended sample (ref: t = -1)"
#| fig-height: 4.5
es_ext <- extract_es(mA_ext)
ggplot(es_ext, aes(x = et, y = est)) +
  geom_hline(yintercept = 0, color = "grey50") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.2, color = "#16a085") +
  geom_point(size = 2.5, color = "#16a085") +
  scale_x_continuous(breaks = -5:5,
                     labels = function(x) ifelse(x == 0, "Recall\nyear",
                                                  as.character(x))) +
  labs(title = "Model A – extended-coverage event-study",
       subtitle = "Agency-city + year FE. SE clustered by city. 95% CIs. Reference t = -1. Endpoints binned at ±5.",
       x = "Event time (years from first successful recall)",
       y = "Coefficient on log(transit capital + 1)")
ggsave(here("output", "figures", "fig_eventstudy_A_extended.png"),
       width = 9, height = 4.5, dpi = 150)
```

```{r}
#| label: 8d-sample-stats
panel |>
  filter(in_modelA_ext) |>
  summarise(
    rows         = n(),
    cities       = n_distinct(city_id),
    treated_cities = n_distinct(city_id[group_A == "treated"]),
    control_cities = n_distinct(city_id[group_A == "control"]),
    agency_city_pairs = n_distinct(agency_city)
  ) |>
  kable(caption = "Extended-coverage Model A sample composition") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Extended-coverage Model B

Same specs as the headline Model B on the extended-coverage sample
(`in_modelB_ext`): baseline and `× covid_period`.

```{r}
#| label: 8e-fit
dat_mB_ext   <- panel |> filter(in_modelB_ext, !is.na(log_transit_cap),
                                !is.na(log_pop_density), !is.na(unemp_rate))
mB_ext       <- fit_modelB_static(dat_mB_ext)
mB_ext_covid <- fit_modelB_covid(dat_mB_ext)

etable(mB_ext, mB_ext_covid,
       headers = c("Ext", "Ext × COVID"),
       title   = "Model B – extended-coverage sample",
       digits  = "r3", se.below = TRUE)
```

```{r}
#| label: 8e-sample-stats
dat_mB_ext |>
  summarise(
    rows              = n(),
    cities            = n_distinct(city_id),
    agency_city_pairs = n_distinct(agency_city)
  ) |>
  kable(caption = "Extended-coverage Model B regression sample") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Restricted-coverage Model A

Restricted sample (`in_modelA_rest`): (city, agency) pairs that are both
HQ-matched and listed in the FTA RY2024 service-area data (SA-place or
SA-county).

```{r}
#| label: 8f-fit
dat_mA_rest     <- panel |> filter(in_modelA_rest, !is.na(log_transit_cap))
dat_mA_rest_ctl <- dat_mA_rest |>
  filter(!is.na(log_pop_density), !is.na(unemp_rate))

mA_rest_nc <- feols(
  log_transit_cap ~ i(et_for_reg, ref = -1)
  | agency_city + year, cluster = ~ city_id, data = dat_mA_rest)

mA_rest <- fit_modelA(dat_mA_rest_ctl)

etable(mA_rest_nc, mA_rest,
       headers = c("No ctrl", "Baseline"),
       title   = "Model A – restricted-coverage sample (HQ ∩ area-served)",
       digits  = "r3", se.below = TRUE)
```

```{r}
#| label: 8f-plot
#| fig-cap: "Model A event-study coefficients with 95% CIs – restricted sample (ref: t = -1)"
#| fig-height: 4.5
es_rest <- extract_es(mA_rest)
ggplot(es_rest, aes(x = et, y = est)) +
  geom_hline(yintercept = 0, color = "grey50") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.2, color = "#8e44ad") +
  geom_point(size = 2.5, color = "#8e44ad") +
  scale_x_continuous(breaks = -5:5,
                     labels = function(x) ifelse(x == 0, "Recall\nyear",
                                                  as.character(x))) +
  labs(title = "Model A – restricted-coverage event-study",
       subtitle = "Agency-city + year FE. SE clustered by city. 95% CIs. Reference t = -1. Endpoints binned at ±5.",
       x = "Event time (years from first successful recall)",
       y = "Coefficient on log(transit capital + 1)")
ggsave(here("output", "figures", "fig_eventstudy_A_restricted.png"),
       width = 9, height = 4.5, dpi = 150)
```

```{r}
#| label: 8f-sample-stats
dat_mA_rest_ctl |>
  summarise(
    rows              = n(),
    cities            = n_distinct(city_id),
    treated_cities    = n_distinct(city_id[group_A == "treated"]),
    control_cities    = n_distinct(city_id[group_A == "control"]),
    agency_city_pairs = n_distinct(agency_city)
  ) |>
  kable(caption = "Restricted-coverage Model A sample composition (controlled spec)") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Restricted-coverage Model B

Same restricted sample applied to Model B (`in_modelB_rest`): no-controls,
with-controls, and `× covid_period` variants.

```{r}
#| label: 8g-fit
dat_mB_rest     <- panel |> filter(in_modelB_rest, !is.na(log_transit_cap))
dat_mB_rest_ctl <- dat_mB_rest |>
  filter(!is.na(log_pop_density), !is.na(unemp_rate))

mB_rest_nc <- feols(
  log_transit_cap ~ any_petition_this_year
  | agency_city + year, cluster = ~ city_id, data = dat_mB_rest)

mB_rest       <- fit_modelB_static(dat_mB_rest_ctl)
mB_rest_covid <- fit_modelB_covid(dat_mB_rest_ctl)

etable(mB_rest_nc, mB_rest, mB_rest_covid,
       headers = c("No ctrl", "Baseline", "× COVID"),
       title   = "Model B – restricted-coverage sample (HQ ∩ area-served)",
       digits  = "r3", se.below = TRUE)
```

```{r}
#| label: 8g-sample-stats
dat_mB_rest_ctl |>
  summarise(
    rows              = n(),
    cities            = n_distinct(city_id),
    agency_city_pairs = n_distinct(agency_city)
  ) |>
  kable(caption = "Restricted-coverage Model B regression sample") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Headline summary

```{r}
#| label: tab-headline
bind_rows(
  extract_one(mA_ext,       "Model A extended (t=0)",
              RECALL_T0_PAT, "Recall (t=0)"),
  extract_one(mB_ext,       "Model B extended",
              PETITION_PAT,  "Petition"),
  extract_one(mB_ext_covid, "Model B ext × COVID",
              PETITION_PAT,  "Petition (non-COVID)"),
  extract_one(mB_ext_covid, "Model B ext × COVID",
              COVID_INT_PAT, "Petition × COVID"),
  extract_one(mA_rest,        "Model A restricted (t=0)",
              RECALL_T0_PAT, "Recall (t=0)"),
  extract_one(mB_rest_nc,     "Model B restricted (no ctrl)",
              PETITION_PAT,  "Petition"),
  extract_one(mB_rest,        "Model B restricted",
              PETITION_PAT,  "Petition"),
  extract_one(mB_rest_covid,  "Model B restricted × COVID",
              PETITION_PAT,  "Petition (non-COVID)"),
  extract_one(mB_rest_covid,  "Model B restricted × COVID",
              COVID_INT_PAT, "Petition × COVID")
) |>
  mutate(`95% CI` = sprintf("[%.3f, %.3f]", est - 1.96 * se, est + 1.96 * se),
         across(c(est, se, p), ~ round(.x, 3))) |>
  select(spec, term, est, se, `95% CI`, p) |>
  kable(caption = "Headline coefficients") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Model A coverage cascade (combined table)

The same `fit_modelA` event-study estimated on each of the three coverage tiers, written as a single LaTeX table (one column per sample) for the thesis.

```{r}
#| label: 8h-modelA-cascade
mods_A <- list(
  "EXT"  = fit_modelA(panel |> filter(in_modelA_ext)),
  "HQ"   = fit_modelA(panel |> filter(in_modelA_hq)),
  "REST" = mA_rest
)

# Order the event-time terms across all three fits, then append the controls.
et_terms <- mods_A |>
  lapply(\(m) grep("^et_for_reg::", names(coef(m)), value = TRUE)) |>
  unlist() |> unique()
et_terms <- et_terms[order(as.integer(str_extract(et_terms, "-?\\d+")))]

coef_map_A <- c(setNames(sub("et_for_reg::", "t = ", et_terms), et_terms),
                "log_pop_density" = "log(pop density)",
                "unemp_rate"      = "Unemployment rate")

# On-screen HTML view: pull modelsummary's formatted numbers as a data frame
# and render with kable, so the column headers stay clean.
modelsummary::modelsummary(
  mods_A,
  output   = "data.frame",
  stars    = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = coef_map_A,
  gof_omit = "AIC|BIC|RMSE|Log.Lik"
) |>
  mutate(term = if_else(statistic == "std.error", "", term)) |>
  select(` ` = term, EXT, HQ, REST) |>
  kable(caption = "Model A event-study across coverage tiers (EXT / HQ / REST)") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r}
#| label: 8h-modelA-cascade-tex
modelsummary::modelsummary(
  mods_A,
  output   = here("output", "tables", "modelA_cascade.tex"),
  stars    = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = coef_map_A,
  gof_omit = "AIC|BIC|RMSE|Log.Lik",
  title    = "Model A event-study across coverage tiers",
  notes    = list(
    "Event-study coefficients on log(transit capital + 1). Reference period t = -1. Window endpoints are binned.",
    "Agency-city and year fixed effects. Standard errors clustered by city.",
    "EXT: agency matched by headquarters, service-area place, or service-area county. HQ: agency headquartered in the city. REST: agency both headquartered in and serving the city."
  )
)
cat("Wrote LaTeX Model A cascade to output/tables/modelA_cascade.tex\n")
```

### Pre-trend joint Wald test by tier

Joint F-test that the pre-period event-time coefficients (`et_for_reg::-2` and earlier) are zero, run separately on each coverage tier.

```{r}
#| label: 8h-modelA-pretrend
pre_pat_A <- "^et_for_reg::-[2-9]|^et_for_reg::-[1-9][0-9]"
imap(mods_A, \(m, nm) pretrend_wald(m, pre_pat_A, nm)) |>
  list_rbind() |>
  mutate(across(c(F_stat, p_value), ~ round(.x, 3))) |>
  kable(caption = "Model A pre-trend joint Wald test by tier (H0: all pre-period coefs = 0)",
        col.names = c("Sample", "Pre-period terms", "F", "p")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Model B coverage cascade (combined table)

The static petition DiD estimated on each of the three coverage tiers, written as a single table. Columns 1-3 are the baseline spec (petition dummy + controls); columns 4-6 add the COVID interaction so the petition effect can differ in 2020-2021.

```{r}
#| label: 8i-modelB-cascade-prep
dat_mB_hq2 <- panel |> filter(in_modelB_hq, !is.na(log_transit_cap),
                              !is.na(log_pop_density), !is.na(unemp_rate))

mods_B <- list(
  fit_modelB_static(dat_mB_ext),
  fit_modelB_static(dat_mB_hq2),
  fit_modelB_static(dat_mB_rest_ctl),
  fit_modelB_covid(dat_mB_ext),
  fit_modelB_covid(dat_mB_hq2),
  fit_modelB_covid(dat_mB_rest_ctl)
)
names(mods_B) <- c("EXT", "HQ", "REST", "EXT ", "HQ ", "REST ")

covid_term <- grep(":covid_period", names(coef(mods_B[[4]])), value = TRUE)[1]
coef_map_B <- c(
  "any_petition_this_yearTRUE" = "Petition filed",
  setNames("Petition × COVID", covid_term),
  "log_pop_density"            = "log(pop density)",
  "unemp_rate"                 = "Unemployment rate"
)
```

```{r}
#| label: 8i-modelB-cascade
ms_df <- modelsummary::modelsummary(
  mods_B, output = "data.frame",
  stars    = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = coef_map_B,
  gof_omit = "AIC|BIC|RMSE|Log.Lik")

mod_cols <- tail(names(ms_df), 6)
ms_df |>
  mutate(term = if_else(statistic == "std.error", "", term)) |>
  select(` ` = term, all_of(mod_cols)) |>
  kable(caption = "Model B petition DiD across coverage tiers (EXT / HQ / REST)",
        col.names = c(" ", "EXT", "HQ", "REST", "EXT", "HQ", "REST")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  add_header_above(c(" " = 1, "Baseline" = 3, "× COVID (2020-21)" = 3))
```

```{r}
#| label: 8i-modelB-cascade-tex
tt <- modelsummary::modelsummary(
  mods_B, output = "tinytable",
  stars    = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = coef_map_B,
  gof_omit = "AIC|BIC|RMSE|Log.Lik",
  title    = "Model B petition DiD across coverage tiers",
  notes    = list(
    "Coefficients on log(transit capital + 1). Petition filed = 1 in a petition-filing year.",
    "Columns 1-3: baseline. Columns 4-6: petition interacted with a 2020-2021 indicator.",
    "Agency-city and year fixed effects. Standard errors clustered by city.",
    "EXT: agency matched by headquarters, service-area place, or service-area county. HQ: agency headquartered in the city. REST: agency both headquartered in and serving the city."
  )
)
tt <- tinytable::group_tt(tt, j = list("Baseline" = 2:4, "× COVID (2020-21)" = 5:7))
tinytable::save_tt(tt, output = here("output", "tables", "modelB_cascade.tex"),
                   overwrite = TRUE)
cat("Wrote LaTeX Model B cascade to output/tables/modelB_cascade.tex\n")
```

---

## Model B petition effect across tiers (figure)

The petition effect outside COVID (the petition coefficient) and during COVID
(petition + petition × COVID), with 95% CIs, for each coverage tier. The
during-COVID effect is a linear combination of two coefficients: its estimate
and SE come from `car::deltaMethod` and its p-value from `car::linearHypothesis`
(a Wald test of `b1 + b2 = 0`), both using the model's clustered covariance. The
companion table reports all six points so every plotted estimate, CI, and p is
traceable.

```{r}
#| label: 8j-prep
combo_effect <- function(m, tier) {
  co <- coef(m); ct <- summary(m)$coeftable
  b1 <- "any_petition_this_yearTRUE"
  b2 <- grep(":covid_period", names(co), value = TRUE)[1]
  # during-COVID effect (b1 + b2): estimate and SE from car::deltaMethod, p from
  # car::linearHypothesis (Wald test of b1 + b2 = 0); both use the clustered vcov.
  # outside-COVID is the native petition coefficient (matches the cascade table).
  dm <- car::deltaMethod(m, paste0("`", b1, "` + `", b2, "`"))
  hm <- setNames(rep(0, length(co)), names(co)); hm[c(b1, b2)] <- 1
  lh <- car::linearHypothesis(m, matrix(hm, nrow = 1), rhs = 0)
  tibble(tier   = tier,
         period = c("Outside COVID", "During COVID (2020-21)"),
         est    = c(ct[b1, 1], dm$Estimate),
         se     = c(ct[b1, 2], dm$SE),
         p      = c(ct[b1, 4], lh[2, grep("^Pr", names(lh))]))
}

es_B <- bind_rows(
  combo_effect(fit_modelB_covid(dat_mB_ext),      "EXT"),
  combo_effect(fit_modelB_covid(dat_mB_hq2),      "HQ"),
  combo_effect(fit_modelB_covid(dat_mB_rest_ctl), "REST")
) |>
  mutate(lo = est - 1.96 * se, hi = est + 1.96 * se,
         tier   = factor(tier, levels = c("EXT", "HQ", "REST")),
         period = factor(period, levels = c("Outside COVID", "During COVID (2020-21)")))
```

```{r}
#| label: 8j-modelB-plot
#| fig-cap: "Model B petition effect by tier – outside vs during COVID (95% CIs, ref: no petition)"
#| fig-height: 4.5
ggplot(es_B, aes(x = est, y = fct_rev(tier), color = period)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_pointrange(aes(xmin = lo, xmax = hi),
                  position = position_dodge(width = 0.55), linewidth = 0.7) +
  scale_color_manual(values = c("Outside COVID" = "#c0392b",
                                "During COVID (2020-21)" = "#2980b9")) +
  labs(title = "Model B – petition effect on transit capital by coverage tier",
       subtitle = "Static petition DiD with COVID interaction. Points = effect, bars = 95% CIs.\nAgency-city + year FE. SE clustered by city.",
       x = "Coefficient on log(transit capital + 1)", y = NULL, color = NULL) +
  theme(legend.position = "bottom")

ggsave(here("output", "figures", "fig_modelB_coef.png"),
       width = 9, height = 4.5, dpi = 150)
```

```{r}
#| label: 8j-modelB-margins-table
margins_B <- es_B |>
  arrange(tier, period) |>
  transmute(Tier = as.character(tier),
            Period = as.character(period),
            Effect   = sprintf("%.3f", est),
            SE       = sprintf("%.3f", se),
            `95% CI` = sprintf("[%.3f, %.3f]", lo, hi),
            p        = sprintf("%.3f", p))

margins_B |>
  kable(caption = paste0(
    "Petition effect by period (data behind the figure). ",
    "Outside COVID = the petition coefficient; during COVID = petition + ",
    "petition x COVID, with SE combined via the fitted covariance. ",
    "95% CI uses +/- 1.96 SE; during-COVID p from a linear-combination test.")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

tt_m <- tinytable::tt(
  margins_B,
  caption = "Model B petition effect by period and coverage tier",
  notes   = list(
    "Outside COVID = the petition coefficient (b1); during COVID = b1 + b2 (petition + petition x COVID).",
    "During-COVID SE combines both terms via the fitted covariance. 95% CI uses +/- 1.96 SE.",
    "Outside-COVID p is the petition coefficient's p; during-COVID p is a delta-method linear-combination test of b1 + b2 = 0.",
    "From the COVID-interaction spec on each tier: agency-city and year fixed effects, SE clustered by city."
  ))
tinytable::save_tt(tt_m, output = here("output", "tables", "modelB_margins.tex"),
                   overwrite = TRUE)
cat("Wrote LaTeX Model B margins to output/tables/modelB_margins.tex\n")
```

