---
title: "Stage 9: Robustness checks"
---

```{r}
#| label: purpose
#| include: false
# =============================================================================
# Purpose : Robustness checks for the headline specs (agency-city-year panel).
#           Each check is run on all three samples (EXT / HQ / REST). Model A
#           reports the recall-year (t=0) event-study coefficient; Model B uses
#           the COVID-interaction spec (petition main effect = non-COVID effect).
#   9.1  Pre-trend Wald test (Model A)
#   9.2  Event-window variation (±3, ±5, ±7) for Model A
#   9.3  Alternative DV: NTD operating expenses
#   9.4  Local-government hard filter (keep only general-purpose local-gov agencies)
# Inputs  : data/final/panel.rds
# Outputs : output/figures/, inline HTML
# =============================================================================
```

## Setup

```{r}
#| label: setup
#| cache.extra: !expr file.info(here::here("data", "final", "panel.rds"))$mtime
library(tidyverse)
library(fixest)
library(kableExtra)
library(here)
here::i_am("_quarto.yml")
source(here("R", "model_helpers.R"))

panel <- readRDS(here("data", "final", "panel.rds"))

dir.create(here("output", "tables"),  recursive = TRUE, showWarnings = FALSE)
dir.create(here("output", "figures"), recursive = TRUE, showWarnings = FALSE)

theme_set(theme_minimal(base_size = 12) +
          theme(plot.title = element_text(face = "bold"),
                plot.subtitle = element_text(color = "grey40")))

setFixest_dict(c(
  log_transit_cap          = "log(transit capital + 1)",
  log_transit_op           = "log(transit operating + 1)",
  log_transit_total        = "log(transit total + 1)",
  log_pop_density          = "log(pop density)",
  unemp_rate               = "Unemployment rate (%)",
  any_petition_this_year   = "Petition (any)",
  city_id                  = "City",
  year                     = "Year"
))
```

```{r}
#| label: window-helper
# The panel carries the default ±5 binning; `make_et` overrides the bin
# endpoints for the window-variation robustness check.
make_et <- function(df, lo, hi) {
  df |>
    mutate(
      et_binned  = case_when(
        is.na(event_time_A) ~ NA_integer_,
        event_time_A <= lo  ~ as.integer(lo),
        event_time_A >= hi  ~ as.integer(hi),
        TRUE                ~ as.integer(event_time_A)
      ),
      et_for_reg = ifelse(group_A == "control" | is.na(et_binned), -1L, et_binned)
    )
}
```


## Baseline replications

Agency-city-year panel: FE = agency_city + year; cluster = city_id. Every
robustness check below is run on all three samples (EXT / HQ / REST). Model A
reports the recall-year (t=0) event-study coefficient; Model B is the
COVID-interaction spec, whose petition main effect is the non-COVID effect (the
reported headline) and whose interaction term is the pandemic-year shift.

```{r}
#| label: baselines
# Per-tier Model A event-study fits, used by the pre-trend Wald test below.
mA_ext  <- fit_modelA(panel |> filter(in_modelA_ext))
mA_hq   <- fit_modelA(panel |> filter(in_modelA_hq))
mA_rest <- fit_modelA(panel |> filter(in_modelA_rest))

# Pre-period regex for the Model A pre-trend Wald test (et < -1).
PRE_PAT <- "^et_for_reg::-[2-9]|^et_for_reg::-[1-9][0-9]"
```

---

## Pre-trend Wald test

Joint F-test that Model A pre-period coefficients are zero
(`et_for_reg::-2` through `et_for_reg::-5`).

```{r}
#| label: 9-1-modelA-wald
bind_rows(
  pretrend_wald(mA_ext,  PRE_PAT, "EXT"),
  pretrend_wald(mA_hq,   PRE_PAT, "HQ"),
  pretrend_wald(mA_rest, PRE_PAT, "REST")
) |>
  mutate(across(c(F_stat, p_value), ~ round(.x, 3))) |>
  kable(caption = "Model A – joint Wald test on pre-period event-time coefs (et < -1), by sample") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Event-window variation

Re-fit the Model A event-study at three window widths (±3, ±5 baseline, ±7).

```{r}
#| label: 9-2-modelA-windows
fit_modelA_window <- function(df, lo, hi) {
  df_w <- df |> make_et(lo, hi)
  feols(log_transit_cap ~ i(et_for_reg, ref = -1) + log_pop_density + unemp_rate
        | agency_city + year, cluster = ~ city_id,
        data = df_w |> filter(!is.na(log_transit_cap),
                              !is.na(log_pop_density), !is.na(unemp_rate)))
}

# Three windows on each tier. ±5 is the baseline already in the panel.
windows_for <- function(flag) {
  d <- panel |> filter(.data[[flag]])
  list(`±3` = fit_modelA_window(d, -3, 3),
       `±5` = fit_modelA(d),
       `±7` = fit_modelA_window(d, -7, 7))
}
win_ext  <- windows_for("in_modelA_ext")
win_hq   <- windows_for("in_modelA_hq")
win_rest <- windows_for("in_modelA_rest")

for (nm in c("EXT", "HQ", "REST")) {
  w <- switch(nm, EXT = win_ext, HQ = win_hq, REST = win_rest)
  print(etable(w[["±3"]], w[["±5"]], w[["±7"]],
               headers = c("±3", "±5 (baseline)", "±7"),
               title   = paste0("Model A event-window variation – ", nm, " sample"),
               digits  = "r3", se.below = TRUE))
}
```

```{r}
#| label: 9-2-plot
#| fig-cap: "Model A event-study at three window widths, by sample"
#| fig-height: 4.5
#| fig-width: 11
es_windows <- bind_rows(
  bind_rows(extract_es(win_ext[["±3"]],  "±3"),
            extract_es(win_ext[["±5"]],  "±5 (baseline)"),
            extract_es(win_ext[["±7"]],  "±7")) |> mutate(tier = "EXT"),
  bind_rows(extract_es(win_hq[["±3"]],   "±3"),
            extract_es(win_hq[["±5"]],   "±5 (baseline)"),
            extract_es(win_hq[["±7"]],   "±7")) |> mutate(tier = "HQ"),
  bind_rows(extract_es(win_rest[["±3"]], "±3"),
            extract_es(win_rest[["±5"]], "±5 (baseline)"),
            extract_es(win_rest[["±7"]], "±7")) |> mutate(tier = "REST")
) |>
  mutate(tier = factor(tier, levels = c("EXT", "HQ", "REST")))

ggplot(es_windows, aes(x = et, y = est, color = spec, fill = spec)) +
  geom_hline(yintercept = 0, color = "grey50") +
  geom_vline(xintercept = -0.5, linetype = "dashed", color = "grey40") +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.2,
                position = position_dodge(width = 0.4)) +
  geom_point(size = 2.5, position = position_dodge(width = 0.4)) +
  facet_wrap(~ tier, scales = "free_y") +
  scale_x_continuous(breaks = -7:7) +
  labs(title = "Model A – event-window robustness",
       subtitle = "Coefficients on event-time dummies; reference = -1.",
       x = "Event time (years from first successful recall)",
       y = "Coefficient on log(transit capital + 1)",
       color = "Window", fill = "Window")

ggsave(here("output", "figures", "fig_robust_windows_A.png"),
       width = 11, height = 4.5, dpi = 150)
```

The end-period-width table (below) reports the restricted-sample recall-year
(t=0) and event-year-2 coefficients at each width. Interior event-times only.

```{r}
#| label: tab-robust-A
cellA <- function(m, pat) {
  ct <- summary(m)$coeftable; r <- grep(pat, rownames(ct))
  if (length(r)) sprintf("%.3f (%.3f)", ct[r[1], 1], ct[r[1], 2]) else NA_character_
}
rowA <- function(m, lbl) tibble(
  `End-period width`  = lbl,
  `Recall year (t=0)` = cellA(m, "et_for_reg::0$"),
  `Event-year 2`      = cellA(m, "et_for_reg::2$"))

robust_A <- bind_rows(
  rowA(win_rest[["±5"]], "5 years (baseline)"),
  rowA(win_rest[["±3"]], "3 years"),
  rowA(win_rest[["±7"]], "7 years"))

capA_title <- "Model A end-period-width robustness, restricted sample"
capA_note  <- paste(
  "Event-study coefficients on log(transit capital + 1), city-clustered standard",
  "errors in parentheses, as the end-period width (the cutoff beyond which distant",
  "event-years are pooled into the endpoint indicators) is set to 3, 5 (baseline),",
  "and 7 years.")

writeLines(
  as.character(
    kbl(robust_A, format = "latex", booktabs = TRUE, caption = capA_title) |>
      footnote(general = capA_note, general_title = "",
               footnote_as_chunk = TRUE, threeparttable = TRUE, escape = FALSE)),
  here("output", "tables", "tab_robust_modelA.tex"))

robust_A |>
  kable(caption = capA_title) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  footnote(general = capA_note, general_title = "", footnote_as_chunk = TRUE)
```

---

## Alternative DV: operating expenses

Re-fit the headline specs with operating expenses (`log_transit_op`) as the
dependent variable, on all three samples. This tests whether recall activity
moves the short-term operating budget rather than the capital budget. Model A
reports the t=0 event-study coefficient; Model B is the COVID-interaction spec.

```{r}
#| label: 9-4-alt-dv-fits
fit_modelA_dv <- function(df, dv) {
  fml <- as.formula(sprintf(
    "%s ~ i(et_for_reg, ref = -1) + log_pop_density + unemp_rate | agency_city + year",
    dv))
  feols(fml, cluster = ~ city_id,
        data = df |> filter(!is.na(.data[[dv]]),
                            !is.na(log_pop_density), !is.na(unemp_rate)))
}

fit_modelB_covid_dv <- function(df, dv) {
  fml <- as.formula(sprintf(
    "%s ~ any_petition_this_year * covid_period + log_pop_density + unemp_rate | agency_city + year",
    dv))
  feols(fml, cluster = ~ city_id,
        data = df |> filter(!is.na(.data[[dv]]),
                            !is.na(log_pop_density), !is.na(unemp_rate)))
}

etable(fit_modelA_dv(panel |> filter(in_modelA_ext),  "log_transit_op"),
       fit_modelA_dv(panel |> filter(in_modelA_hq),   "log_transit_op"),
       fit_modelA_dv(panel |> filter(in_modelA_rest), "log_transit_op"),
       headers = c("EXT", "HQ", "REST"),
       title   = "Alt DV log(operating + 1) – Model A event-study, by sample",
       digits  = "r3", se.below = TRUE)

etable(fit_modelB_covid_dv(panel |> filter(in_modelB_ext),  "log_transit_op"),
       fit_modelB_covid_dv(panel |> filter(in_modelB_hq),   "log_transit_op"),
       fit_modelB_covid_dv(panel |> filter(in_modelB_rest), "log_transit_op"),
       headers = c("EXT", "HQ", "REST"),
       title   = "Alt DV log(operating + 1) – Model B (COVID interaction), by sample",
       digits  = "r3", se.below = TRUE)
```

Operating expenses as an alternative outcome, restricted sample, reported as a
full table for each model in parallel with the main results tables: the Model A
event-study, and the Model B petition DiD (baseline and COVID interaction), both
on operating spending.

```{r}
#| label: tab-robust-operating-A
op_A_rest <- fit_modelA_dv(panel |> filter(in_modelA_rest), "log_transit_op")

et_terms_op <- grep("^et_for_reg::", names(coef(op_A_rest)), value = TRUE)
et_terms_op <- et_terms_op[order(as.integer(str_extract(et_terms_op, "-?\\d+")))]
coef_map_op_A <- c(setNames(sub("et_for_reg::", "t = ", et_terms_op), et_terms_op),
                   "log_pop_density" = "log(pop density)",
                   "unemp_rate"      = "Unemployment rate")

modelsummary::modelsummary(
  list("Operating expenses" = op_A_rest),
  output   = here("output", "tables", "tab_robust_operating_A.tex"),
  stars    = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = coef_map_op_A,
  gof_omit = "AIC|BIC|RMSE|Log.Lik",
  title    = "Model A robustness: operating expenses as the outcome (restricted sample)",
  notes    = list(
    "Event-study coefficients on log(transit operating + 1). Reference period t = -1. Distant event-years are pooled into the endpoint indicators.",
    "Agency-city and year fixed effects. Standard errors clustered by city."))
cat("Wrote LaTeX Model A operating table to output/tables/tab_robust_operating_A.tex\n")

modelsummary::modelsummary(
  list("Operating expenses" = op_A_rest), output = "data.frame",
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = coef_map_op_A, gof_omit = "AIC|BIC|RMSE|Log.Lik") |>
  mutate(term = if_else(statistic == "std.error", "", term)) |>
  select(` ` = term, `Operating expenses`) |>
  kable(caption = "Model A operating-expenses event-study (restricted sample)") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r}
#| label: tab-robust-operating-B
fit_modelB_static_dv <- function(df, dv) {
  feols(as.formula(sprintf(
    "%s ~ any_petition_this_year + log_pop_density + unemp_rate | agency_city + year", dv)),
    cluster = ~ city_id,
    data = df |> filter(!is.na(.data[[dv]]), !is.na(log_pop_density), !is.na(unemp_rate)))
}
op_B_rest <- list(
  "Baseline" = fit_modelB_static_dv(panel |> filter(in_modelB_rest), "log_transit_op"),
  "× COVID"  = fit_modelB_covid_dv(panel  |> filter(in_modelB_rest), "log_transit_op"))

covid_term_op <- grep(":covid_period", names(coef(op_B_rest[[2]])), value = TRUE)[1]
coef_map_op_B <- c(
  "any_petition_this_yearTRUE" = "Petition filed",
  setNames("Petition × COVID", covid_term_op),
  "log_pop_density"            = "log(pop density)",
  "unemp_rate"                 = "Unemployment rate")

modelsummary::modelsummary(
  op_B_rest,
  output   = here("output", "tables", "tab_robust_operating_B.tex"),
  stars    = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = coef_map_op_B,
  gof_omit = "AIC|BIC|RMSE|Log.Lik",
  title    = "Model B robustness: operating expenses as the outcome (restricted sample)",
  notes    = list(
    "Coefficients on log(transit operating + 1). Petition filed = 1 in a petition-filing year.",
    "The second column interacts petition with a 2020-2021 indicator.",
    "Agency-city and year fixed effects. Standard errors clustered by city."))
cat("Wrote LaTeX Model B operating table to output/tables/tab_robust_operating_B.tex\n")

modelsummary::modelsummary(
  op_B_rest, output = "data.frame",
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  coef_map = coef_map_op_B, gof_omit = "AIC|BIC|RMSE|Log.Lik") |>
  mutate(term = if_else(statistic == "std.error", "", term)) |>
  select(` ` = term, Baseline, `× COVID`) |>
  kable(caption = "Model B operating-expenses petition DiD (restricted sample)") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

## Local-government hard filter

Re-fits the headline specs after dropping all agency-city rows where the agency
is not controlled by a general-purpose local government (`local_controlled ==
FALSE`). Keeps only the "City, County or Local Government Unit or DOT" category;
drops Independent Public Agencies, State DOTs, MPOs, Tribes, private providers,
universities, and non-profits.

```{r}
#| label: 9-5-prep
panel_local <- panel |> filter(local_controlled)
cat(sprintf("Locally-controlled-only panel: %d rows / %d cities (vs full %d / %d)\n",
            nrow(panel_local), n_distinct(panel_local$city_id),
            nrow(panel), n_distinct(panel$city_id)))

panel |> distinct(ntd_id, local_controlled) |>
  count(local_controlled) |>
  kable(caption = "Agencies by local_controlled flag") |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r}
#| label: 9-5-fits
etable(fit_modelA(panel_local |> filter(in_modelA_ext)),
       fit_modelA(panel_local |> filter(in_modelA_hq)),
       fit_modelA(panel_local |> filter(in_modelA_rest)),
       headers = c("EXT", "HQ", "REST"),
       title   = "Local-government hard filter – Model A event-study, by sample",
       digits  = "r3", se.below = TRUE)

etable(fit_modelB_covid(panel_local |> filter(in_modelB_ext)),
       fit_modelB_covid(panel_local |> filter(in_modelB_hq)),
       fit_modelB_covid(panel_local |> filter(in_modelB_rest)),
       headers = c("EXT", "HQ", "REST"),
       title   = "Local-government hard filter – Model B (COVID interaction), by sample",
       digits  = "r3", se.below = TRUE)
```

The local-government table reports the restricted-sample non-COVID petition
coefficient under the local-government subset, against the baseline. The cities
column reports the number of cities remaining under each restriction.

```{r}
#| label: tab-robust-B
rest_B <- panel |> filter(in_modelB_rest)

petB <- function(m) {
  ct <- summary(m)$coeftable
  r  <- grep("^any_petition_this_year(?:TRUE)?$", rownames(ct), perl = TRUE)
  sprintf("%.3f (%.3f)", ct[r[1], 1], ct[r[1], 2])
}
nC <- function(df) df |>
  filter(!is.na(log_transit_cap), !is.na(log_pop_density), !is.na(unemp_rate)) |>
  pull(city_id) |> n_distinct()

robust_B <- tibble(
  Specification = c("Baseline (restricted sample)",
                    "Local-government agencies only"),
  `Petition (non-COVID)` = c(
    petB(fit_modelB_covid(rest_B)),
    petB(fit_modelB_covid(rest_B |> filter(local_controlled)))),
  Cities = c(nC(rest_B), nC(rest_B |> filter(local_controlled))))

capB_title <- "Model B local-government robustness, restricted sample"
capB_note  <- paste(
  "Non-COVID petition coefficient (the petition main effect in the",
  "COVID-interaction specification) on log(transit capital + 1), city-clustered",
  "standard errors in parentheses. The second row keeps only agencies that are an",
  "arm of a local government.")

writeLines(
  as.character(
    kbl(robust_B, format = "latex", booktabs = TRUE, caption = capB_title) |>
      footnote(general = capB_note, general_title = "",
               footnote_as_chunk = TRUE, threeparttable = TRUE, escape = FALSE)),
  here("output", "tables", "tab_robust_modelB.tex"))

robust_B |>
  kable(caption = capB_title) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  footnote(general = capB_note, general_title = "", footnote_as_chunk = TRUE)
```

